---
title: "Basic inferential analysis of tooth growth data"
---

# Basic inferential analysis of tooth growth data

## Synopsis

We will perform basic inferential analysis for the length of teeth in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

## Data Load

Let's load data:

```{r}
library(datasets)
data(ToothGrowth)
head(ToothGrowth)
```

## Exploratory analysys

Data size:

```{r}
dim(ToothGrowth)
```

Summary of the data:

```{r}
summary(ToothGrowth)
```

Some exploratory plots:

```{r}
library(ggplot2)
ggplot(aes(x = dose, y = len), data = ToothGrowth) + 
geom_point(aes(color = supp)) +
labs(x = 'Dose in milligrams') +
labs(y = 'Tooth length') +
labs(title = 'Tooth growth for each of dose levels of Vitamin C with each of methods')
```

```{r}
ggplot(aes(x = supp, y = len), data = ToothGrowth) +
geom_boxplot(aes(fill = supp)) + 
facet_wrap(~ dose) +
labs(y = 'Tooth length') +
labs(title = 'Tooth growth for each of dose levels of Vitamin C with each of methods')
```

## Confidence Intervals and Hypothesis Testing

Data is neither paired nor with equal variance. 

Let's perform hypothesis testing for different circumstances:

### Dosage only

Firsly let's split data for paired comparison of dose levels (we will compare only nearest pairs):

```{r}
ToothGrowth_05vs10 <- subset(ToothGrowth, dose %in% c(0.5, 1.0))
ToothGrowth_10vs20 <- subset(ToothGrowth, dose %in% c(1.0, 2.0))
```

```{r}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = ToothGrowth_05vs10)
```

```{r}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = ToothGrowth_10vs20)
```

Both tests brings the same conclusion: bigger dose increases tooth growth.

### Supplement type for different doses

Let's now test supplement types for different doses:

```{r}
ToothGrowth_05 <- subset(ToothGrowth, dose == 0.5)
ToothGrowth_10 <- subset(ToothGrowth, dose == 1.0)
ToothGrowth_20 <- subset(ToothGrowth, dose == 2.0)
```

```{r}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = ToothGrowth_05)
```

Conclusion: for 0.5mg of Vitamin C, orange juice is better than ascorbic acid for tooth growth.

```{r}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = ToothGrowth_10)
```

Conclusion: for 1mg of Vitamin C, orange juice is better than ascorbic acid for tooth growth.

```{r}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = ToothGrowth_20)
```

Conclusion: for 2mg of Vitamin C, there is no difference between orange juice or ascorbic acid.

## Conclusions

To sum up gotten conclusions:

* Bigger dose increases tooth growth.
* For 0.5mg and 1mg of Vitamin C, orange juice is better than ascorbic acid for tooth growth.
* For 2mg of Vitamin C, there is no difference between orange juice or ascorbic acid.